---
title: "EggRatio_analyses2025"
output: html_document
date: "2025-02-03"
---

This markdown file contains analyses for the Egg Count field data collected in 2022. Raw egg count was previously cleaned to correct spelling errors and erroneous dates. Prevalence data was similarly cleaned and processed prior to analyses. 

First, load all necessary libraries

```{r, load_libraries}
#libraries
library(ggplot2)
library(dplyr)
library(emmeans)
library(glmmTMB)
library(remotes)
library(DHARMa)
library(ggpubr)
library(tidyr)
library(pscl)
library(boot)
library(MASS)
library(car)
#install.packages("boot")
#install.packages("Matrix", version="1.6-2")
#install.packages("DHARMa")
#install.packages("remotes")
#install.packages("ggpubr")
#install.packages("pscl")
```
I am working with egg count data saved as cleaned_eggcounts_includesephippial_1_29_26_tas.csv. 
```{r, data_read}
#cleaned egg count data
all_data <- read.csv(file="Data/cleaned_eggcounts_includesephippial_1_29_26_TAS.csv", header=T)
head(all_data)
```

Remove ephippia & observations past sexual selection for cerio & dentifera.

```{r, ephippia_and_ephippiadates}
#Epphipia are included in these data. Since we are doing a comparison of egg counts, we won't be using those observations. However, since the start of sexual reproduction might also skew the frequency of eggs in general, I am going to remove all observations after the onset of sexual reproduction for dentifera & cerio in the fall. 

#Prior to removing all ephippia, use theme to determine the likely onset of sexual reproduction in the populations. 
dent <- all_data[all_data$Host.Species=="Dentifera"&all_data$Egg.Number=="e",]
cerio <- all_data[all_data$Host.Species=="Cerio"&all_data$Egg.Number=="e",]

#identify all unique julian dates for ephippial dent and cerio
unique(dent$julian)
unique(cerio$julian)
#identified the dates that sexual reproduction started: 285 for both cerio and dentifera
#note: there is an single ephippia at day 158 in dentifera, but I am not going to treat that as the start date

#filter out these dates
df <- all_data[!(all_data$Host.Species %in% c("Dentifera", "Cerio") & all_data$julian >= 285), ]
  #filtering in this way removes 1528 observations

#now to remove remaining ephippia. 
#create a logical vector for which all non-ephippial individuals are given a 1
LV_noepphipial <- df$Egg.Number!="e"
sum(LV_noepphipial) #sum is 14861, meaning there are 61 ephippia observations to remove

df <-df[LV_noepphipial,] 
#new data frame has 14861 observations, which matches the removal of 61 observations (the remaining ephippial animals)

#double check Egg Number column to ensure that all ephippia have been removed
class(df$Egg.Number)
unique(df$Egg.Number)

#reclassify as numeric
df$Egg.Number<- as.numeric(df$Egg.Number)

#make sure julian is a factor, not an integer
df$julian <- as.numeric(df$julian)

#check that we didn't filter ALL the large dates out
unique(df$julian)


#check dates for manuscript reasons
#unique(df$Date.Sampled)
```
Now I'll filter out the rare parasites.

```{r, filter_rare_parasites}
#figure out which parasites we won't be looking at
table(df$Parasite)
#output: 
#Brood  Caullerya       Coke    Gurleya  Larssonia     Metsch       MicG       Past     Spider      Spiro Uninfected 
 #       57         30       2030         10          4         12       1050        337          1        152      11178 
#get rid of Caullerya, gurleya, larssonia, metsch, spider

#filter out the parasites of lowest density
df_filtered <- df[df$Parasite=="Brood"|df$Parasite=="Past"|df$Parasite=="Coke"|df$Parasite=="MicG"|df$Parasite=="Uninfected",]

#check that filtering worked
unique(df_filtered$Parasite)

#add in a month column so that models can be simplified later on
#set date column to date class
df_filtered$Date.Sampled <- as.Date(df_filtered$Date.Sampled)

#extract month
df_filtered$month <- format(df_filtered$Date.Sampled, "%m")
unique(df_filtered$month)
```
Coke is Pan. perplexa
Brood is B. paedophthorum
Past is Pas. ramosa
MicG is O. pajunii

"df_filtered" contain all data from observations of uninfected animals as well as 5 parasites which had more than 50 observations over the course of the 2022 field season.
Create separate dataframes for each of the parasites. Eventually, I will join together the uninfected data with data for each of the parasites, subsetting the uninfected so that only data from dates and lakes where the parasite was found are kept. 

```{r, subset_data}
#first, split out the original data frame by parasite
brood_only <- df_filtered[df_filtered$Parasite=="Brood",]
past_only <- df_filtered[df_filtered$Parasite=="Past",]
micg_only <-df_filtered[df_filtered$Parasite=="MicG",]
coke_only <- df_filtered[df_filtered$Parasite=="Coke",]
uninf_only <- df_filtered[df_filtered$Parasite=="Uninfected",]
```

Now I need to try properly subsetting the data so that the uninfected animals I'm comparing to infected are ONLY from the lakes, dates and hosts where the infected were found. This is trickier than it seems because if we subset using only logical vectors,then we run into the problem where hosts where infections were not found but were in lakes on a date where brood was found in a different host will still be hanging around. 
Start with brood since it has the least observations 


```{r, data_wrangling_brood_test}

#merge brood and uninfected
brood_test <- rbind(brood_only, uninf_only) #currently contains all the uninfected & brood observations

#concatenate host species, lake & date 
brood_test$specieslakeday <- paste(brood_test$Host.Species,brood_test$Lake.Name,brood_test$julian)

#now we just need to match specieslakedates of the infected animals with the same for uninfected
#subset data by whether or not animal is infected
Ybrood <- dplyr::filter(brood_test, Parasite=="Brood")
Nbrood <- dplyr::filter(brood_test, Parasite=="Uninfected")

#look in negative dataset to see if the specieslakeday is in the positive data set
# if it is, test variable will be yes
Nbrood$test <-  ifelse(Nbrood$specieslakeday %in% Ybrood$specieslakeday, "yes", "no")

#create the test variable in the data set with all infected animals
Ybrood$test <- "yes"

#recombine data sets so it's brood and uninfected animals
recombinedbrood <- rbind (Ybrood, Nbrood)

#subset data by the test variable being yes
brood_proper <- dplyr::filter(recombinedbrood, test=="yes") #final data set has 851 observations
```

Repeat the above with four other parasites (spiro, past, micg, sprite)

```{r, subsetting_properly}
################################################################################################
####MICG######
#merge micg and uninfected
micg_test <- rbind(micg_only, uninf_only) #currently contains all the uninfected & micg observations

#concatenate host species, lake & date 
micg_test$specieslakeday <- paste(micg_test$Host.Species,micg_test$Lake.Name,micg_test$julian)

#now we just need to match specieslakedates of the infected animals with the same for uninfected
#subset data by whether or not animal is infected
Ymicg <- dplyr::filter(micg_test, Parasite=="MicG")
Nmicg <- dplyr::filter(micg_test, Parasite=="Uninfected")

#look in negative dataset to see if the specieslakeday is in the positive data set
# if it is, test variable will be yes
Nmicg$test <-  ifelse(Nmicg$specieslakeday %in% Ymicg$specieslakeday, "yes", "no")

#create the test variable in the data set with all infected animals
Ymicg$test <- "yes"

#recombine data sets so it's micg and uninfected animals
recombinedmicg <- rbind (Ymicg, Nmicg)

#subset data by the test variable being yes
micg_proper <- dplyr::filter(recombinedmicg, test=="yes") #final data set has 4041 observations

##########################################################################################
######Coke
#merge coke and uninfected
coke_test <- rbind(coke_only, uninf_only) #currently contains all the uninfected & coke observations

#concatenate host species, lake & date 
coke_test$specieslakeday <- paste(coke_test$Host.Species,coke_test$Lake.Name,coke_test$julian)

#now we just need to match specieslakedates of the infected animals with the same for uninfected
#subset data by whether or not animal is infected
Ycoke <- dplyr::filter(coke_test, Parasite=="Coke")
Ncoke <- dplyr::filter(coke_test, Parasite=="Uninfected")

#look in negative dataset to see if the specieslakeday is in the positive data set
# if it is, test variable will be yes
Ncoke$test <-  ifelse(Ncoke$specieslakeday %in% Ycoke$specieslakeday, "yes", "no")

#create the test variable in the data set with all infected animals
Ycoke$test <- "yes"

#recombine data sets so it's coke and uninfected animals
recombinedcoke <- rbind (Ycoke, Ncoke)

#subset data by the test variable being yes
coke_proper <- dplyr::filter(recombinedcoke, test=="yes") #final data set has 7067 observations

##########################################################################################
######Past
#merge past and uninfected
past_test <- rbind(past_only, uninf_only) #currently contains all the uninfected & past observations

#concatenate host species, lake & date 
past_test$specieslakeday <- paste(past_test$Host.Species,past_test$Lake.Name,past_test$julian)

#now we just need to match specieslakedates of the infected animals with the same for uninfected
#subset data by whether or not animal is infected
Ypast <- dplyr::filter(past_test, Parasite=="Past")
Npast <- dplyr::filter(past_test, Parasite=="Uninfected")

#look in negative dataset to see if the specieslakeday is in the positive data set
# if it is, test variable will be yes
Npast$test <-  ifelse(Npast$specieslakeday %in% Ypast$specieslakeday, "yes", "no")

#create the test variable in the data set with all infected animals
Ypast$test <- "yes"

#recombine data sets so it's past and uninfected animals
recombinedpast <- rbind (Ypast, Npast)

#subset data by the test variable being yes
past_proper <- dplyr::filter(recombinedpast, test=="yes") #final data set has 2084 observations
```

rename merged dataframes

```{r, rename_datasets}
brood <-brood_proper
past <- past_proper
micg <- micg_proper
coke <- coke_proper

```
___________________________________________________________
___________________________________________________________
________________Table 1____________________________________
___________________________________________________________
___________________________________________________________
Descriptive stats/counts for table 1
```{r}
table(coke_only$Host.Species)
table(micg_only$Host.Species)
table(past_only$Host.Species)
table(brood_only$Host.Species)

table(coke_only$Lake.Name)
table(micg_only$Lake.Name)
table(past_only$Lake.Name)
table(brood_only$Lake.Name)

prev <- read.csv(file="data/Prev_data_eggratio_2022.csv", header=T)

max(prev$coke.prev, na.rm=TRUE)
max(prev$pasteuria.prev, na.rm=TRUE)
max(prev$micG.prev, na.rm=TRUE)
max(prev$brood.prev, na.rm=TRUE)


```
Counts/prevalence across individual host species

```{r}

#table things
table(uninf_only$Host.Species, uninf_only$Lake.Name)
table(uninf_only$Host.Species)

table(coke_only$Host.Species,coke_only$Lake.Name)
table(coke_only$Host.Species)


prev %>%
  dplyr::group_by(Host.Species) %>%
  dplyr::summarise(max_value = max(coke.prev, na.rm = TRUE), .groups = "drop")

prev_parv <-prev[prev$Host.Species=="parvula",]
max(prev_parv$coke.prev, na.rm=TRUE)


table(micg_only$Host.Species,micg_only$Lake.Name)
table(micg_only$Host.Species)

prev %>%
  dplyr::group_by(Host.Species) %>%
  dplyr::summarise(max_value = max(micG.prev, na.rm = TRUE), .groups = "drop")

table(past_only$Host.Species,past_only$Lake.Name)
table(past_only$Host.Species)

prev %>%
  dplyr::group_by(Host.Species) %>%
  dplyr::summarise(max_value = max(pasteuria.prev, na.rm = TRUE), .groups = "drop")

table(brood_only$Host.Species,brood_only$Lake.Name)
table(brood_only$Host.Species)
prev %>%
  dplyr::group_by(Host.Species) %>%
  dplyr::summarise(max_value = max(brood.prev, na.rm = TRUE), .groups = "drop")

```

___________________________________________________________
___________________________________________________________
________________STATS-MODELS____________________________
___________________________________________________________
___________________________________________________________

These data are zero-inflated. Zeroes can represent a true biological "zero" where the individual animal did not have any babies or a "false" zero where the individual had just had babies so their brood chamber was empty. Because of these factors, I think the best way to proceed with analyses is a ZIP (zero-inflated poisson). This is similar in some ways to a hurdle model, except in a ZIP model both groups (both zeroes and non-zeroes) are considered in separate models. 

```{r, pick_terms}
#get a load of this zero-inflation (coke-only dataset)
hist(coke$Egg.Number)
table(coke$Egg.Number)

#we know it's zero-inflated, but what terms should we be including? Are lake/date necessary? I am also collapsing date here into month since I want to simplify the models as much as possible
#look at lake & month 
boxplot(Egg.Number ~ Lake.Name, data = df_filtered)
boxplot(Egg.Number ~ month, data = df_filtered)

#do rough comparison 
kruskal.test(Egg.Number ~ Lake.Name, data = df_filtered)
kruskal.test(Egg.Number ~ month, data = df_filtered)
  #both above tests are significant across all data, indicating that lake & month should be included in models where possible. 
```

####################PAN PERPLEXA ################################


First, do preliminary checks to ensure that ZIP/ZINB is appropriate
```{r, model_checks_coke}
#Check Outcome variable (Egg Number)
table(coke$Egg.Number==0)
mean(coke$Egg.Number==0) #41% of values are zeroes, making a zero-inflated model necessary

#Check for overdispersion
mean(coke$Egg.Number) #1.094948
var(coke$Egg.Number)  #1.494097
  #variance is higher than the mean, meaning overdispersion is present and negative binomial is preferred over poisson

#fit both models and compare goodness-of-fit
poisson_fit <- glm(Egg.Number ~ Parasite+Host.Species, data=coke, family="poisson")
nb_fit <- glm.nb(Egg.Number ~ Parasite+Host.Species, data=coke)
AIC(poisson_fit, nb_fit) #Negative binomial has a lower AIC, it is the more appropriate model. 


#fit basic poisson & zero-inflated versions then compare with Vuong test
zip_fit <- zeroinfl(Egg.Number ~ Parasite+Host.Species | Parasite+Host.Species,data=coke, dist="poisson")
zinb_fit <- zeroinfl(Egg.Number ~ Parasite+Host.Species | Parasite+Host.Species,data=coke, dist="negbin")
vuong(zip_fit, poisson_fit) # ZIP vs Poisson
vuong(zinb_fit, nb_fit)     # ZINB vs NB
```
These tests show that I need a zero-inflated model, and that negative binomial is likely more appropriate than a poisson distribution.

Ok, having tried a bunch of models there are a ton of difficulties with these data. I still think conceptually ZIP/ZINB models are the best way to go, but these models do not want to converge, especially when I try to include Lake & month as random OR fixed effects. 

Difficulties with converging are likely are result of unbalanced data across levels (escpeially lake), where host species are rare/absent in some lakes but not others. 

Below are all attempted/run models for pansporella, running models to optimize the inclusion of random effects and/or a zero-inflation term. 
1- What I think is the "ideal" model: ZINB with parasite & host species in both models including lake & month as random effects in both count & zero models. 
2- ZINB model with lake only parasite & host species in both count & zero models
3- ZINB model with fixed predictor in the zero-inflation term (functionally, this is assuming that any single individual is equally likely to have a structural or sampling zero)
5-Standard GLM with Negative binomial or poisson distribution on untransformed data



Now, before trying all these alternative models, I am going to filter out the rare host species for this parasite. I am doing this to simplify the model terms so I can get models to converge. 

```{r}
#table host species to determine which ones I am excluding
#first, look at which hosts are the least abundant overall
table(coke$Host.Species)
#now look at host and parasite (are there any hosts with very few infections?)
table(coke$Host.Species, coke$Parasite)
  #lowest count hosts are cerio & mendotae
  #NOTE: tried model with only cerio & mendotae excluded but it still failed to converge with all necessary terms included
  #Try excluding Dubia & ambigua as well (both hosts with less than 100 infections)

#make vector of species I am keeping for analyses
keep_species <- c("Dubia", "Ambigua","Dentifera","Parvula","Pulicaria","Retrocurva")

#filter out rare host species
coke_filt <- coke %>% 
  filter(Host.Species %in% keep_species)

table(coke_filt$Host.Species)
```
Try all models

```{r, ZIP_coke}

#1a-What I think is the "ideal" model: ZINB with parasite & host species in both models including lake & month as random effects in both count & zero models. 
#zinb_mod_1a <- glmmTMB(
#  Egg.Number ~ Host.Species+ Parasite+ Host.Species * Parasite + (1|Lake.Name)+(1|month),
#  ziformula = ~ Host.Species + Parasite +Host.Species * Parasite + (1|Lake.Name)+(1|month),       # predictors for zero-inflation
#  family = nbinom2,
#  data = coke_filt
#)
#this model does not converge unfortunately. After looking into each of this variables individually in test models it seems like a large part of the issue is that some of these factors have "deterministic" zero behavior; in other words, some lakes either always or never have zeroes. This makes the zero-inflation term fail. 
#If deterministic zero behavior makes a ZINB fail, that also tells me that  I probably won't be able to use this kind of model for parasites where I expect deterministic zeroes for the infection status. In other words, if I expect past & brood-infected animals to nearly always have zeroes (which I know in these data they do), any zero-inflated model that I use will most likely fail. 

#1b
#zinb_mod_1b <- glmmTMB(
#  Egg.Number ~ Host.Species * Parasite+ (1|month),
#  ziformula = ~ Host.Species + Parasite,       # predictors for zero-inflation
#  family = nbinom2,
#  data = coke_filt
#)
#this model converges!
#However, it only converges if I take the individual main effects out of the count model and keep the interaction (or keep the individual main effects and exclude the interaction). The later I think is allowed but does not address my question precisely where the former I don't believe is allowed? I don't think you should ever have your interaction term in a model without the individual main effects?
#Again, when looking at lakes it seems like the major problem is that some lakes are always or never zero; I can potentially collapse lakes to decrease levels but I am not sure the most biologically relevant way to go about that. 
# I will look at the model summary & all the checks/comparisons later down
#plot residuals with DHARMa
#sim_res_1b <- simulateResiduals(zinb_mod_1b)
#plot(sim_res_1b) #only KS test significant
#testDispersion(sim_res_1b) #non-significant
#testZeroInflation(sim_res_1b) #ns


#2-ZINB model with lake only parasite & host species in both count & zero models
zinb_mod_2 <- glmmTMB(
  Egg.Number ~ Host.Species+Parasite+Host.Species:Parasite  ,
  ziformula = ~ Host.Species + Parasite+Host.Species:Parasite,       # predictors for zero-inflation
  family = nbinom2,
  data = coke_filt
)
sim_res_2 <- simulateResiduals(zinb_mod_2)
plot(sim_res_2) #outlier test significant, residual vs. predicted looks better than 1b
testDispersion(sim_res_2) #non-significant
testZeroInflation(sim_res_2) #ns

#this model converges. Again, though, looking at all the data indicate that lake & month are likely important and should be included.



#3- ZINB model with fixed predictor in the zero-inflation term 
zinb_mod_3 <- glmmTMB(
  Egg.Number ~Host.Species+Parasite+ Host.Species:Parasite + (1|month),
  ziformula = ~ 1,       # predictors for zero-inflation
  family = nbinom2,
  data = coke_filt
)


sim_res_3 <- simulateResiduals(zinb_mod_3)
plot(sim_res_3) 
testDispersion(sim_res_3) #non-significant
testZeroInflation(sim_res_3) #ns


#5-Standard GLM with Negative binomial or poisson distribution on untransformed data
zinb_mod_5 <- glmmTMB(
  Egg.Number ~ Host.Species + Parasite +Host.Species:Parasite +(1|Lake.Name)+(1|month), 
  family=nbinom2,
  data = coke_filt
)
#test model
sim_res_5 <- simulateResiduals(zinb_mod_5)
plot(sim_res_5) #KS test significant, residual vs. predicted somewhat wobbly
testDispersion(sim_res_5) #non-significant
testZeroInflation(sim_res_5) #ns
#this test shows non-significant zero inflation

zinb_mod_5b <- glmmTMB(
  Egg.Number ~ Host.Species + Parasite+ Host.Species * Parasite +(1|Lake.Name/julian), 
  family=poisson,
  data = coke_filt
)
sim_res_5b <- simulateResiduals(zinb_mod_5b)
plot(sim_res_5b) #same results as sim_res_5 (above)
testDispersion(sim_res_5b) #non-significant
testZeroInflation(sim_res_5b) #ns

#try negative binomial distribution

zinb_mod_5c <- glmmTMB(
  Egg.Number ~ Host.Species + Parasite+ Host.Species * Parasite +(1|Lake.Name/julian), 
  family=nbinom1,
  data = coke_filt
)
sim_res_5c <- simulateResiduals(zinb_mod_5c)
plot(sim_res_5c) #same results as sim_res_5 (above)
testDispersion(sim_res_5c) #non-significant
testZeroInflation(sim_res_5c) #
#compare AICs
AIC(zinb_mod_2)
AIC(zinb_mod_3)
AIC(zinb_mod_5)
AIC(zinb_mod_5b)
AIC(zinb_mod_5c)
#model 5 has the best fit of the model 5 options. Note that I am not comparing AICs of model 2 and 3 because they include different terms and AIC is therefore an inappropriate compariative metric

#Based on above information, it is important to choose the most biologically relevant/logical model
```

Look at predictions output by each model and compare to see if there are significant differences before selecting a final model.

```{r}
#compare models by simulating residuals  (this is a copy of above )
res2 <- simulateResiduals(zinb_mod_2, n = 500)
res5 <- simulateResiduals(zinb_mod_5, n = 500)
res3 <- simulateResiduals(zinb_mod_3, n = 500)

plot(res2); testZeroInflation(res2); testDispersion(res2)
plot(res5); testZeroInflation(res5); testDispersion(res5)
plot(res3); testZeroInflation(res3); testDispersion(res3)

#plot model predictions to predict what is happening 
prediction_2 <- predict(zinb_mod_2, type = "response")
plot(prediction_2)
predictions_5 <- predict(zinb_mod_5, type = "response")
predictions_3 <- predict(zinb_mod_3, type = "response")
predictions_5b <- predict(zinb_mod_5b, type="response")


#use predict to plot model
nd <- expand.grid(
  Host.Species = sort(unique(coke_filt$Host.Species)),
  Parasite     = sort(unique(coke_filt$Parasite)),
  KEEP.OUT.ATTRS = FALSE,
  stringsAsFactors = FALSE
)

nd$pred_mod2 <- predict(zinb_mod_2, newdata = nd, type = "response")
nd$pred_mod5 <- predict(zinb_mod_5, newdata = nd, type = "response", re.form = ~0)
nd$pred_mod3 <- predict(zinb_mod_3, newdata = nd, type = "response", re.form = ~0)
nd$pred_mod5b <- predict(zinb_mod_5b, newdata = nd, type = "response", re.form = ~0)


nd$zi_mod2 <- predict(zinb_mod_2, newdata = nd, type = "zprob")
nd$zi_mod3 <- predict(zinb_mod_3, newdata = nd, type = "zprob", re.form = ~0)

ggplot(nd, aes(x = Parasite, y = pred_mod3, group = Host.Species, color = Host.Species)) +
  geom_point(position = position_dodge(width = 0.2)) +
  geom_line(position = position_dodge(width = 0.2)) +
  labs(y = "Predicted mean egg count", title = "Predictions from zinb_mod_3") +
  theme_bw()



nd_long <- tidyr::pivot_longer(
  dplyr::select(nd, Host.Species, Parasite, pred_mod2, pred_mod3, pred_mod5, pred_mod5b),
  cols = dplyr::starts_with("pred_"),
  names_to = "model",
  values_to = "pred"
)
ggplot(nd_long, aes(x = Parasite, y = pred, color = Host.Species, group = Host.Species)) +
  geom_point() +
  geom_line() +
  facet_wrap(~ model) +
  labs(y = "Predicted mean egg count", title = "Model predictions by Host species and infection status") +
  theme_bw()




#add in confidence intervals

get_emm_ci <- function(fit, model_name){
  out <- emmeans(fit, ~ Parasite | Host.Species, type="response") |>
    as.data.frame()

  pred_name <- intersect(c("response","rate","prob","emmean"), names(out))[1]
  out$response <- out[[pred_name]]

  out |> mutate(model = model_name)
}

emm2 <- get_emm_ci(zinb_mod_2, "mod_2 (ZI, no RE)")
emm3 <- get_emm_ci(zinb_mod_3, "mod_3 (ZI + month RE)")
emm5 <- get_emm_ci(zinb_mod_5, "mod_5 (lake+month RE, no ZI)")
emm5b <- get_emm_ci(zinb_mod_5b, "mod_5b (lake+month RE, Lake/time)")


emm_all <- bind_rows(emm2, emm3, emm5, emm5b)
emm_all$Parasite <- factor(emm_all$Parasite, levels = c("Uninfected","Coke"))


pd <- position_dodge(width = 0.25)

ggplot(emm_all, aes(x = Parasite, y = response,
                    color = Host.Species, group = Host.Species)) +
  geom_point(position = pd) +
  geom_line(position = pd) +
  geom_errorbar(aes(ymin = asymp.LCL, ymax = asymp.UCL),
                width = 0.15, position = pd) +
  facet_wrap(~ model) +
  labs(x = "Infection status",
       y = "Predicted mean egg count (95% CI)") +
  theme_bw()



#looking at these prediction outputs, the magnitude of response seems to vary widely across models, but responses themselves are largely similar 
```

Run outputs for each possible model and pairwise comparisons 

Note: Model 3 was selected as the model we would use from this point forward. Outputs of the tests below did not contribute to this decision; we decided on this model because it was the most biologically relevant stable model which was able to converge given difficulties with these data. Tests below are run on all the potential models to compare outputs as "sanity checks". 
```{r}
#look at summary of mod_3 and mod_5
summary(zinb_mod_3)
summary(zinb_mod_5)
summary(zinb_mod_5b)

#Likelihood ratio test to see if infection x host is an important consideration
m3_full <- zinb_mod_3
m3_red  <- update(m3_full, . ~ . - Host.Species:Parasite)

anova(m3_red, m3_full) #both significant



m5_full <- zinb_mod_5
m5_red  <- update(m5_full, . ~ . - Host.Species:Parasite)

anova(m5_red, m5_full) #both significant 



m5b_full <- zinb_mod_5b
m5b_red  <- update(m5b_full, . ~ . - Host.Species:Parasite)

anova(m5b_red, m5b_full) #both significant 


#look at the contributions of random effects to varability 
VarCorr(m5_full)
VarCorr(m5b_full)
VarCorr(m3_full)



#Do pairwise comparisons 
emm5 <- emmeans(m5_full, ~ Parasite | Host.Species, type = "response")
pairs(emm5, adjust = "tukey")  # or "holm" / "fdr"
#ambig, dent, parv, retro all significant

emm3 <- emmeans(m3_full, ~ Parasite | Host.Species, type = "response")
pairs(emm3, adjust = "tukey")
#ambig, dent, retro significant
#ambigua has an unreasonably high response ratio.

emm5b <- emmeans(m5b_full, ~ Parasite | Host.Species, type = "response")
pairs(emm5b, adjust = "tukey")
#only ambig, dent, retro significant



# contrasts (infected vs uninfected within each species)
pw <- pairs(emmeans(zinb_mod_3, ~ Parasite | Host.Species, type = "response"))


# 95% CIs for the ratios
confint(pw, level = 0.95)
```

```{r, ambigua_coke_check}

#pairwise comparisons for pansporella model return very high ratio for Ambigua
pairs(emm3, adjust = "tukey") #this should be diagnosed to determine cause and decide if this pairwise comparison should be excluded from the manuscript

#Likely causes of particularly large ratio estimates are either zero-inflation in ambigua itself or unbalanced numbers in the groups being compaired. D. ambigua are the smallest group being compared so that might cause the high ratio estimate

table(coke$Host.Species, coke$Parasite) #only 32 infected ambigua, compared to 194 uninfected. This is unbalanced, but not much more than dubia 

#graph only ambigua
ggplot(subset(coke, Host.Species == "Ambigua"),
       aes(x = Parasite, y = Egg.Number)) +
  geom_boxplot(outlier.alpha = 0.3) +
  geom_jitter(width = 0.1, alpha = 0.4)


#create an official subset to look at specific data points
ambigua_coke <- subset(coke, Host.Species=="Ambigua")

table(ambigua_coke$Parasite, ambigua_coke$Egg.Number)

#from this plot and table, it seems like the issue could be the low N combined with the presence of an some outlier (1 counts of 4)


#quantify influence on ratio by comparing model which excludes that point

# identify row with outlier
idx <- c(329)

fit0 <- m3_full
fit_drop <- update(m3_full, data = coke_filt[-idx, ])

# compare the problematic contrast (example: Host within InfStatus)
emm0 <- emmeans(fit0, ~ Parasite | Host.Species, type="response")
emm1 <- emmeans(fit_drop, ~ Parasite | Host.Species, type="response")

pairs(emm0, adjust="tukey")
pairs(emm1, adjust="tukey")

#ratio is significantly improved by the exclusion of that point, but there is no reason to believe that the point is not a legitimate egg count, and should therefore not be actually excluded. Another contributing factor might be zero inflation in the uninfected ambigua; and zero is by far the most common egg count in ambigua in the coke dataframe

#determine contribution of zero inflation in ambigua specifically
sim_all <- simulateResiduals(m3_full)
testZeroInflation(sim_all)

# subset to Ambigua only
sim_amb <- DHARMa::recalculateResiduals(sim_all, sel=coke_filt$Host.Species == "Ambigua")
testZeroInflation(sim_amb)
#zero-inflation in ambigua term is significant, but model term cannot support more robust zero-inflation term

#we will keep the model as is, and the pairwise comparisons, but we will not provide any interpretation of the ambigua pairwise comparison in the table. 
```






####################PAS. RAMOSA MODELS################################

```{r}
#table host species to determine which ones I am excluding
#first, look at which hosts are the least abundant overall
table(past$Host.Species)
#now look at host and parasite (are there any hosts with very few infections?)
table(past$Host.Species, past$Parasite)
  #For this model I will keep all host species
```
The number of possible models for past is lower because it has deterministic zeroes. That means that any model that has a ZI term will not be able to converge because infection status will nearly always have 100% zeroes. 

```{r}
#model 1: all terms with lake and julian nested, try different distributions 
past_mod_1<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name/julian), family = nbinom1(link="log"), data = past)

past_mod_1_res <- simulateResiduals(past_mod_1, n = 500)
plot(past_mod_1_res); testZeroInflation(past_mod_1_res); testDispersion(past_mod_1_res)
#fit looks good

#negbinom2
past_mod_2<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name/julian), family = nbinom2(link="log"), data = past)

past_mod_2_res <- simulateResiduals(past_mod_2, n = 500)
plot(past_mod_2_res); testZeroInflation(past_mod_2_res); testDispersion(past_mod_2_res)

#poisson
past_mod_3<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name/julian), family = poisson, data = past)

past_mod_3_res <- simulateResiduals(past_mod_3, n = 500)
plot(past_mod_3_res); testZeroInflation(past_mod_3_res); testDispersion(past_mod_3_res)
#within group deviations form uniformity significant

#poisson
past_mod_4<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name)+(1|month), family = poisson, data = past)


past_mod_4_res <- simulateResiduals(past_mod_4, n = 500)
plot(past_mod_4_res); testZeroInflation(past_mod_4_res); testDispersion(past_mod_4_res)


#nbinom1
past_mod_5<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name)+(1|month), family = nbinom1, data = past)

past_mod_5_res <- simulateResiduals(past_mod_5, n = 500)
plot(past_mod_5_res); testZeroInflation(past_mod_5_res); testDispersion(past_mod_5_res)


AIC(past_mod_1, past_mod_2, past_mod_3, past_mod_4, past_mod_5) 
#model 2 does not produce an AIC, and both other models have stunningly similar AICs


#model 1 is what we are moving foward with
```

Look at past model output
```{r}
#look at summary of mod_3 and mod_5
summary(past_mod_1)

#LRT 
past_full <- past_mod_1
past_red  <- update(past_full, . ~ . - Host.Species:Parasite)

anova(past_full, past_red)
#anova is not significant WHICH IS FINE. Here is just means there is not a significant interaction of infection x host. Which we expect for past

####
car::Anova(past_full, type = 3)  # or type=2 depending on your preference
#only parasite is significant, which again we largely expect


VarCorr(past_full) #lake name is 0.04 which is fairly low




library(emmeans)

emm_past <- emmeans(past_full, ~ Parasite | Host.Species, type = "response")
pairs(emm_past, adjust = "tukey")  # or "holm" / "fdr"
#dent, parv, retro all significant

# contrasts (infected vs uninfected within each species)
pw_past <- pairs(emmeans(past_full, ~ Parasite | Host.Species, type = "response"))


# 95% CIs for the ratios
confint(pw_past, level = 0.95)

```


####################B. PAEDOPHTHORUM MODELS################################

```{r}
#table host species to determine which ones I am excluding
#first, look at which hosts are the least abundant overall
table(brood$Host.Species)
#now look at host and parasite (are there any hosts with very few infections?)
table(brood$Host.Species, brood$Parasite)
  #remove ambigua, mendotae, parv

#make vector of species I am keeping for analyses
keep_species <- c("Dentifera","Dubia","Retrocurva", "Cerio", "Pulicaria")

#filter out rare host species
brood_filt <- brood %>% 
  filter(Host.Species %in% keep_species)

table(brood_filt$Host.Species)
```
Models

```{r}
#only poisson distribution for below model converges
brood_mod_1<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name/julian), family = poisson(link="log"), data = brood_filt)


brood_mod_1_res <- simulateResiduals(brood_mod_1, n = 500)
plot(brood_mod_1_res); testZeroInflation(brood_mod_1_res); testDispersion(brood_mod_1_res)
#fit issues: quantile deviations, combined adjusted quantile test significant



#try only lake, not nested lake and julian
brood_mod_1a<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name), family = poisson(link="log"), data = brood_filt)


brood_mod_1a_res <- simulateResiduals(brood_mod_1a, n = 500)
plot(brood_mod_1a_res); testZeroInflation(brood_mod_1a_res); testDispersion(brood_mod_1a_res)
#model still has the same fit errors 

#try negative binomial
brood_mod_1b<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species  + (1|Lake.Name), family = nbinom1(link="log"), data = brood_filt)


brood_mod_1b_res <- simulateResiduals(brood_mod_1b, n = 500)
plot(brood_mod_1b_res); testZeroInflation(brood_mod_1b_res); testDispersion(brood_mod_1b_res)
#quantile deviations still present

#try excluding terms and check fit
VarCorr(brood_mod_1) #should not exclude either of these random effects they both seem pretty important (range 0.36-0.97)


#try removing random effects
brood_mod_2<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species, family = poisson(link="log"), data = brood_filt)

brood_mod_2_res <- simulateResiduals(brood_mod_2, n = 500)
plot(brood_mod_2_res); testZeroInflation(brood_mod_2_res); testDispersion(brood_mod_2_res)
#zero-inflated and over dispersed story of my life


#try different distribution
brood_mod_3<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species, family = nbinom1(link="log"), data = brood_filt)


brood_mod_3_res <- simulateResiduals(brood_mod_3, n = 500)
plot(brood_mod_3_res); testZeroInflation(brood_mod_3_res); testDispersion(brood_mod_3_res)
#no issues!  This makes this the model to beat, but I am excluding random effects

#try adding in month
#brood_mod_4<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species+(1|month), family = nbinom1(link="log"), data = brood_filt)
#does not converge


#remove RE, nbionm2
#brood_mod_4<- glmmTMB(Egg.Number ~ Parasite + Host.Species + Parasite:Host.Species, family = nbinom2(link="log"), data = brood_filt)
#model does not converge

#AIC 
AIC(brood_mod_1, brood_mod_2, brood_mod_3) #model 1 has the best fit, but has significant fit issues. We will move forward with model 3.

``` 

LRT, Anova, Varcorr

```{r}
#look at summary of mod_3 and mod_5
summary(brood_mod_3)
Anova(brood_mod_3)

#LRT 
brood_full <- brood_mod_3
brood_red  <- update(brood_full, . ~ . - Host.Species:Parasite)

anova(brood_full, brood_red)
#anova is not significant WHICH IS FINE. Here is just means there is not a significant interaction of infection x host. Which we expect for brood

####
car::Anova(brood_full, type = 3)  # or type=2 depending on your preference
#only parasite is significant, which again we largely expect


#pairwise comparisons

emm_brood <- emmeans(brood_full, ~ Parasite | Host.Species, type = "response")
pairs(emm_brood, adjust = "tukey")  # or "holm" / "fdr"
#dent, dubia, retro all significant

# contrasts (infected vs uninfected within each species)
pw_brood <- pairs(emmeans(brood_full, ~ Parasite | Host.Species, type = "response"))


# 95% CIs for the ratios
confint(pw_brood, level = 0.95)  #CIs indicate severe issues with pairwise comparisons; model is not robust enough for pairwise comparisons, do not include in manuscript.

```


####################O. PAJUNII MODELS################################

```{r}
#table host species to determine which ones I am excluding
#first, look at which hosts are the least abundant overall
table(micg$Host.Species)
#now look at host and parasite (are there any hosts with very few infections?)
table(micg$Host.Species, micg$Parasite)
  #lowest count hosts are dubia & parv
  #Try excluding Dubia & ambigua as well (both hosts with less than 100 infections)
#after trying out the models below none are converge (even if they have historically listen I am not sure if the packages updated or something). LOoking at the number of infected animals there are only 17 infected pulicaria; remove those as well. 

#make vector of species I am keeping for analyses
keep_species <- c("Cerio","Dentifera","Mendotae","Pulicaria", "Retrocurva")

#filter out rare host species
micg_filt <- micg %>% 
  filter(Host.Species %in% keep_species)

table(micg_filt$Host.Species)
```


Test all alternate models

```{r, micg model testing}
#1a-What I think is the "ideal" model: ZINB with parasite & host species in both models including lake & month as random effects in both count & zero models. 
#zinb_mod_1a <- glmmTMB(
#  Egg.Number ~ Host.Species+ Parasite+ Host.Species * Parasite + (1|Lake.Name)+(1|month),
#  ziformula = ~ Host.Species + Parasite +Host.Species * Parasite + (1|Lake.Name)+(1|month),       # predictors for zero-inflation
#  family = nbinom2,
#  data = micg_filt
#)
#as expected, this model does not converge; there are likely too many factors with either deterministic zeroes or not enough statisical power within the group


#2-ZINB model with lake only parasite & host species in both count & zero models
#JUDAS MODEL BETRAYING ME WORKING AND THEN NOT WORKING 
micg_mod_1 <- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite,
  ziformula = ~ Host.Species + Parasite+Host.Species*Parasite,       # predictors for zero-inflation
  family = nbinom2 (link="log"),
  data = micg_filt
)


micg_mod_1_res <- simulateResiduals(micg_mod_1, n = 500)
plot(micg_mod_1_res); testZeroInflation(micg_mod_1_res); testDispersion(micg_mod_1_res)
#zero-inflated and over dispersed


#try different distributions on the same model
micg_mod_1a <- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite,
  ziformula = ~ Host.Species + Parasite+Host.Species*Parasite,       # predictors for zero-inflation
  family = poisson,
  data = micg_filt
)


micg_mod_1a_res <- simulateResiduals(micg_mod_1a, n = 500)
plot(micg_mod_1a_res); testZeroInflation(micg_mod_1a_res); testDispersion(micg_mod_1a_res)
#same issues; overdispersed and zero-inflated 


#try nbinom1
micg_mod_1b <- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite,
  ziformula = ~ Host.Species + Parasite+Host.Species*Parasite,       # predictors for zero-inflation
  family = nbinom1,
  data = micg_filt
)


micg_mod_1b_res <- simulateResiduals(micg_mod_1b, n = 500)
plot(micg_mod_1b_res); testZeroInflation(micg_mod_1b_res); testDispersion(micg_mod_1b_res)
#same issues; overdispersed and zero-inflated 


#this zero-inflation term is perpetually returning zero-inflated and over dispersed so it's clear the zero-inflation term isn't doing all that much


#3- ZINB model with fixed predictor in the zero-inflation term 
micg_mod_2 <- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite+(1|month),
  ziformula = ~ 1,       # predictors for zero-inflation
  family = nbinom2,
  data = micg_filt
)



micg_mod_2_res <- simulateResiduals(micg_mod_2, n = 500)
plot(micg_mod_2_res); testZeroInflation(micg_mod_2_res); testDispersion(micg_mod_2_res)
#this fit is much better; no longer overdispersed or zero-inflated, and I am able to include a random effect 


#3 modified; try different distributions
micg_mod_2a<- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite+(1|month),
  ziformula = ~ 1,       # predictors for zero-inflation
  family = nbinom1,
  data = micg_filt
)



micg_mod_2a_res <- simulateResiduals(micg_mod_2a, n = 500)
plot(micg_mod_2a_res); testZeroInflation(micg_mod_2a_res); testDispersion(micg_mod_2a_res)
#no issues; this model looks largely the same as the other nbinom model  (mod_2)

#try poisson distribution
micg_mod_2b<- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite+(1|month),
  ziformula = ~ 1,       # predictors for zero-inflation
  family = poisson,
  data = micg_filt
)



micg_mod_2b_res <- simulateResiduals(micg_mod_2b, n = 500)
plot(micg_mod_2b_res); testZeroInflation(micg_mod_2b_res); testDispersion(micg_mod_2b_res)
#this is also a solid model
AIC(micg_mod_2,micg_mod_2a, micg_mod_2b) #poisson has a SLIGHTLY lower AIC 


#Try including nested lake/day in model 
micg_mod_2c<- glmmTMB(
  Egg.Number ~ Host.Species * Parasite + Host.Species+Parasite+(1|Lake.Name/julian),
  ziformula = ~ 1,       # predictors for zero-inflation
  family = poisson,
  data = micg_filt
)
micg_mod_2c_res <- simulateResiduals(micg_mod_2c, n = 500)
plot(micg_mod_2c_res); testZeroInflation(micg_mod_2c_res); testDispersion(micg_mod_2c_res)
#over dispersed 



#-No ZI term, nest lake/julian
micg_mod_4<- glmmTMB(
  Egg.Number ~ Host.Species + Parasite +Host.Species:Parasite+(1|Lake.Name/julian), 
  family=poisson,
  data = micg_filt
)
micg_mod_4_res <- simulateResiduals(micg_mod_4, n = 500)
plot(micg_mod_4_res); testZeroInflation(micg_mod_4_res); testDispersion(micg_mod_4_res)
#Over dispersed

#try different distribution
#micg_mod_4a<- glmmTMB(
 # Egg.Number ~ Host.Species + Parasite +Host.Species:Parasite+(1|Lake.Name/julian), 
 # family=nbinom1,
 # data = micg_filt
#)
#does not converge

#use model 2b
```



```{r}
#look at summary of mod_2b
summary(micg_mod_2b)
Anova(micg_mod_2b)

#LRT 
micg_full <- micg_mod_2b
micg_red  <- update(micg_full, . ~ . - Host.Species:Parasite)

anova(micg_full, micg_red) 
#anova is significant, indicating that host x parasite is important

####
car::Anova(micg_full, type = 3)  # or type=2 depending on your preference
#all terms are significant


#pairwise comparisons

emm_micg <- emmeans(micg_full, ~ Parasite | Host.Species, type = "response")
pairs(emm_micg, adjust = "tukey")  # or "holm" / "fdr"
#dent, dubia, retro all significant

# contrasts (infected vs uninfected within each species)
pw_micg <- pairs(emmeans(micg_full, ~ Parasite | Host.Species, type = "response"))


# 95% CIs for the ratios
confint(pw_micg, level = 0.95)
```


